---
title: "001 Time series"
output: html_notebook
---

The purpose of this notebook is to evaluate the time series of quotations from the prototype version of America's Public Bible and create some visualizations of them.

```{r setup}
library(tidyverse)
library(odbc)
library(TSclust)
library(stringr)
library(RcppRoll)
library(viridis)
library(broman)

con <- dbConnect(odbc::odbc(), "Research DB")
quotations <- tbl(con, "apb_proto_quotations") %>% collect()
verses_by_year <- tbl(con, "apb_proto_verses_by_year") %>%  collect() 
wordcounts <- tbl(con, "apb_proto_wordcounts") %>% collect()
```

What is the difference between the full `quotations` table and the `verses_by_year` table?

```{r}
quotations_count <- quotations %>% 
  group_by(year, reference) %>% 
  summarize(full_count = n()) %>% 
  mutate(reference = str_remove(reference, " \\(KJV\\)")) %>% 
  left_join(verses_by_year, by = c("reference", "year")) %>% 
  arrange(reference, year)
```

By joining them it becomes apparent that the verses by year is a subset of the data. How small a subset?

```{r}
quotations_count_totals <- quotations_count %>% 
  group_by(reference) %>% 
  summarize(n = sum(n, na.rm = TRUE),
            full_count = sum(full_count, na.rm = TRUE)) %>% 
  arrange(desc(n))
quotations_count_totals %>% head(10)
```

The verses by year have the cleaned data, for example, Luke 18:16 includes the other verses in the Synoptics. But it also has fewer quotations in places, probably because it uses higher probability verses.

```{r}
range(quotations_count_totals %>% filter(n > 0) %>% pull(n))
range(quotations_count_totals$full_count)
range(quotations$year)
range(verses_by_year$year)
range(quotations$probability)
```

We are going to use the direct quotations table.

Now we are going to aggregate the verse counts by year and get the word and page counts as well.

```{r}
quotations_per_year <- quotations %>% 
  mutate(reference = str_remove(reference, "\\s\\(KJV\\)")) %>% 
  count(reference, year, testament) %>% 
  group_by(reference) %>% 
  mutate(n_all = sum(n)) %>% 
  left_join(wordcounts, by = "year") %>% 
  mutate(n_per_100mil = n / wordcount * 1e8) %>% 
  arrange(desc(n_all), reference, year) 
quotations_per_year %>% head(10) 
```

Now we can try to plot the time series for the most frequently quoted verses.

```{r}
top_verses <- quotations_per_year %>% 
  ungroup() %>% 
  distinct(reference, n_all) %>% 
  top_n(24, n_all)

verse_order <- top_verses %>% pull(reference) 

# See broman::plot_crayons()
set.seed(97)
col_pal <- brocolors("crayons")[c("Orange Red", "Outrageous Orange", "Brown",
                                  "Beaver", "Raw Sienna", "Raw Umber", "Sunglow",
                                  "Goldenrod", "Asparagus", "Fern", "Green", 
                                  "Pine Green", "Teal Blue", "Pacific Blue", 
                                  "Cerulean", "Green Blue", "Purple Heart", 
                                  "Vivid Violet", "Hot Magenta", "Maroon",
                                  "Scarlet", "Olive Green", "Blush", "Plum")] %>% 
  sample()
names(col_pal) <- verse_order

# Smooth the charts by taking the mean
quotations_per_year %>% 
  ungroup() %>% 
  inner_join(select(top_verses, reference), by = "reference") %>% 
  mutate(reference = ordered(reference, verse_order)) %>% 
  group_by(reference) %>% 
  arrange(reference, year) %>% 
  mutate(smoothed_rate = roll_mean(n_per_100mil, n = 3, fill = NA_real_)) %>% 
  filter(!is.na(smoothed_rate)) %>% 
  ggplot(aes(x = year, y = smoothed_rate, color = reference)) +
  geom_line() + 
  facet_wrap(~reference, ncol = 6) + 
  coord_cartesian(ylim = c(0, 20)) +
  guides(color = FALSE) +
  theme_void() +
  scale_color_manual(values = col_pal) +
  labs(title = "Rates of quotation for popular verses, 1837–1921",
       subtitle = "Top 24 most quoted verses from the KJV in Chronicling America")
```

