---
title: "America's Public Bible: Biblical Quotations in U.S. Newspapers"
---

<style>
  .toc-ignore { display: none;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

<div class="row">
  <div class="col-sm-8 col-xs-12">
  <h1 id="site-title">America's Public Bible</h1>
  <h2 id="site-subtitle">Biblical Quotations in U.S. Newspapers</h2>
  
For most of its issues in 1902, the *Ellensburg Dawn* featured a quotation from Benjamin Franklin prominently on its front page. "A Bible and a newspaper in every house," the quotation proclaimed, "are the principal support of virtue, morality, and civil liberty."^[*Ellensburg Dawn* (Ellensburg, Wash.), [18 Jan. 1902](http://chroniclingamerica.loc.gov/lccn/sn88085012/1902-01-18/ed-1/seq-1/). Also quoted in the *Liberty Advocate* (Liberty, Miss.), [18 Oct. 1845](http://chroniclingamerica.loc.gov/lccn/sn83016942/1845-10-18/ed-1/seq-1/>http://chroniclingamerica.loc.gov/lccn/sn83016942/1845-10-18/ed-1/seq-1/#date1=1836&index=0&rows=20&words=Bible+Franklin+newspaper&searchType=basic&sequence=0&state=&date2=1922&proxtext=bible+newspaper+franklin&y=0&x=0&dateFilterType=yearRange&page=1).] Though the quotation from Franklin was doubtless spurious, the pairing of newspapers with the Bible was common. Throughout the nineteenth century and into the twentieth century, newspapers in the United States---even those newspaper which were not published by a religious denomination or organization---had frequent recourse to the Bible. Newspapers printed sermons and Sunday school lessons, and ministers offered lessons through newspaper Bible clubs.^[*Perrysburg Journal* (Perrysburg, Ohio), [18 Sept. 1919](http://chroniclingamerica.loc.gov/lccn/sn87076843/1919-09-18/ed-1/seq-3/); *Amador Ledger* (Jackson, Amador County, Calif.), [19 Feb. 1909](http://chroniclingamerica.loc.gov/lccn/sn93052980/1909-02-19/ed-1/seq-3/).] They featured jokes whose punchlines required biblical literacy. They printed political commentary that cited the Bible on all sides of a given issue.  They made money off of [advertisements](http://chroniclingamerica.loc.gov/lccn/sn83045433/1918-02-10/ed-1/seq-18/) for Bibles of every kind. Some newspapers even sold Bibles as a way of raising revenue, as when the *[Brattleboro Daily Reformer](http://chroniclingamerica.loc.gov/lccn/sn86071593/1916-11-15/ed-1/seq-6/)* and other papers sold an *Illimunated Holy Bible*. But most of all, newspapers quoted the Bible.

*America's Public Bible* uncovers the presence of biblical quotations in the nearly 11 million newspaper pages in the Library of Congress's *[Chronicling America](http://chroniclingamerica.loc.gov/)* collection. Using the [techniques of machine learning](methods.html) I have identified over 872,000 quotations of or verbal allusions to specific biblical verses on those newspaper pages. For now, the project has looked only for quotations to the [King James Version](https://en.wikipedia.org/wiki/King_James_Version) (or Authorized Version) of the English Bible, by far the most commonly used Bible during the nineteenth-century among American Protestants.^[The American Standard Version of the Bible (the American version of the British Revised Version) was published in 1900. See Peter J. Thuesen, *In Discordance with the Scriptures: American Protestant Battles Over Translating the Bible* (New York: Oxford University Press, 1999).] 

  </div>
  <div class="col-sm-4 col-xs-12">
  
  <div id="splash-image">
  ![An advertisement in the *Brattleboro Daily Reformer* (Brattleboro, Vt.), [15 Nov. 1916](http://chroniclingamerica.loc.gov/lccn/sn86071593/1916-11-15/ed-1/seq-6/).](img/bible-hand.png)
</div>

  </div>
  
</div>

<div class="row">
  
  <div class="col-sm-4 col-xs-12">

  ![This article retells the parable of the good Samaritan (Luke 10:30--36), reinterpreting it with the Populist movement as the Samaritan. From the *Owingsville Outlook* (Owingsville, Ky.), [2 March 1899](http://chroniclingamerica.loc.gov/lccn/sn86069620/1899-03-02/ed-1/seq-4/#words=chance+came+certain+priest+way+saw+side).](img/good-samaritan.png)
  
  </div>
  
  <div class="col-sm-8 col-xs-12">
  
That the Bible was ubiquitous in the nineteenth-century United States is a commonplace claim, but this project aims to provide far more specificity about how the Bible was used in American newspapers. While it is true that the Bible as a whole was nearly everywhere, what matters more are the ways that the Bible was put to use. The Bible offered thousands of texts, each of which could be interpreted in countless ways. For each verse that was quoted, this site offers [a way to explore](exploration.html) both the trend in how frequently that verse was used, and a link to each instance in which it was quoted at *Chronicling America* so that the context of the quotation is available. In this was, the many specifics of how the Bible was used can be understood. The assumption of this project is that the Bible as a text was not the same thing as the Bible as it was put to use by Americans. That is why I have called this America's *public* Bible. By looking at uses of the Bible in newspapers, we can see which parts of the Bible were available to Americans.  Verses that could be cited without a reference could indicate a kind of literacy. This is particularly the case with jokes. Verses that could be cited without explanation indicate a shared assumption or interpretation. Verses that were used constantly speak to a kind of a shared cultural touchstone, while verses that were used episodically reveal the tensions in a particular political or social situation. By looking at how the verses were actually used, we can see how the Bible was a contested yet common text.

The general trend is that the frequency of quotations declined over time among the newspapers in *Chronicling America*, which are not papers published by religious denominations or benevolence organizations.^[I have normalized the number of quotations per year by the number of pages in *Chronicling America* for that year. In examining these charts of general trends a few factors should be kept in mind. First, there are many the number of pages in *Chronicling America* grows over time. Thus, the uncertainty in the trend is greater for the 1840s and 1850s, while it is more certain by the 1910s and 1920s. That said, the peak of quotations in the 1840s makes sense given the widespread expansion of evangelical Protestantism in the 1830s and 1840s. Second, the quality of the optical character recognition (i.e., the process by which the page images of the newspapers are turned into plain text) in the dataset is very uneven: some pages have nearly perfect transcriptions, while others have no usable text. It is of course much more difficult to find quotations, especially brief quotations, in poor OCR. So I cannot claim that I have found every quotation or allusion. Third, a classification problem like this one is always a tradeoff between finding every possible genuine match while avoiding false positives. I have prefered to avoid false positives as much as possible, even at the cost of missing genuine matches that would otherwise have been detected. For more details on how the quotations were found, see the [sources and methods section](methods.html).] ] The period of the 1830s and 1840s saw the expansion of evangelical Protestantism, so that peak represents a time when the Bible was likely to be quoted more frequently. While the Bible was quoted less frequently over time, there were notable resurgences in its use in the 1870s and the 1890s. Quotations of the Bible were never so high that one could find a quotation on every page or every issue, but  The Old Testament and the New Testament followed more or less the same pattern, though the New Testament was always quoted more often than the Old Testament. (Keep in mind, though, that the Old Testament is over three times as long as the New Testament.) 

  </div>
  
</div>

```{r trends-for-whole-bible}
source("helpers.R")
library(dplyr)

bible_by_year <- readRDS("_data/bible-by-year.rds")

bible_by_year %>% 
  plot_ref_ts(title = "Quotations from the Bible in Chronicling America")
```

The general trends, however, tell us much less than the patterns for individual verses. Consider each of the following verses, each of which has a pattern which is strinkingly different than the general trend.^[The trends for individual verses sometimes reflect the reprinting of texts. On reprinting of texts, see See Ryan Cordell, David Smith, et al., *[Viral Texts Mapping Networks of Reprinting in 19th-Century Newspapers and Magazines](http://viraltexts.org/)* (2012--16).]

```{r trends-for-selected-verses}
verses_by_year <- readRDS("_data/verses-by-year.rds")
verses_by_year %>% 
  references_to_ts(c("Luke 18:16", "Proverbs 14:34", "Acts 17:26", "John 15:13")) %>% 
  plot_ref_ts(title = "Trends for selected verses", roll = 3)
```

The most popular verse in these newspapers was "Suffer little children to come unto me, and forbid them not: for of such is the kingdom of God" (Luke 18:16; cf. Matthew 19:14, Mark 10:14). The American Sunday school movement made this verse popular, as organizations like the American Sunday School Union and denominational Sunday schools published Sunday school lessons weekly in the newspapers. This verse peaked in the 1850s, during a formative period for the Sunday school movement.^[Anne M. Boylan, *Sunday School: The Formation of an American Institution, 1790-1880* (New Haven: Yale University Press, 1988).] The text "[God] hath made of one blood all nations of men" (Acts 17:26) was a crucial text with which American Christians thought about race and slavery, especially since it was a proof text for the unity of human beings regardless of race, and since it spoke to scientific debates about monogenesis versus polygenesis (the quetion of whether have a single or multiple origin). This verse peaked in the frequency with which it was quoted in the middle of the Civil War.^[On the Bible and the Civil War, see Mark A. Noll, *The Civil War as a Theological Crisis* (Chapel Hill: University of North Carolina Press, 2006).] While scholars have often discussed the origins of Christian nationalism, newspapers quoted the more common nationalistic verses relatively infrequently compared to other texts. An exception was the text "Righteousness exalteth a nation: but sin is a reproach to any people" (Proverbs 14:34), which was the subject of fast day sermons and other jeremiads during the Civil War.^[Andrew R. Murphy, *Prodigal Nation: Moral Decline and Divine Punishment from New England to 9/11* (New York: Oxford University Press, 2009), 44-76.] While many verses declined in popularity over time, one verse that grew in popularity at the beginning of the twentieth century was John 15:13: "Greater love hath no man than this, that a man lay down his life for his friends." This verse was frequently used to memorialize the dead, especially as associated with heroism. While the verse was used after the Civil War and for victims of a yellow fever epidemic in the Mississippi River valley and Memphis in 1879--80, it peaked when memorializing the dead of the Great War.^[Philip Jenkins, *The Great and Holy War: How World War I Became a Religious Crusade* (San Francisco: HarperOne, 2014).]  Finally, 1 Samuel 3:4 was scarcely ever quoted in American newspapers, with the exception of 1876, when presidential candidate Samuel Tilden adopted "The Lord called Samuel" as his motto.

<div class="row">
  <div class="col-sm-4">
  
  ![The *New York Tribune* identified itself as Republican and anti-slavery by quoting Acts 17:26. Other newspapers reprinted this notice. From the *Lamoille Newsdealer* (Hyde Park, Vt.), [18 Dec. 1862](http://chroniclingamerica.loc.gov/lccn/sn84023428/1862-12-18/ed-1/seq-3/#words=hath+made+blood+nations+men+dwell+face+earth+hath+determined+times+before+appointed+bounds+habitation).](img/ny-tribune.png)
  
  </div>
  <div class="col-sm-4">
  
  ![Democratic newspapers quoted 1 Samuel 3:4 supporting Samuel J. Tilden against Rutherford B. Hayes in the 1876 presidential election. From the *Stark County Democrat* (Canton, Ohio), [09 Nov. 1876](http://chroniclingamerica.loc.gov/lccn/sn84028490/1876-11-09/ed-1/seq-4/#date1=1836&sort=relevance&rows=20&words=called+Lord+Samuel&searchType=basic&sequence=0&index=7&state=&date2=1922&proxtext=%22lord+called+samuel%22&y=0&x=0&dateFilterType=yearRange&page=2).](img/lord-called-samuel.png)
  
  </div>
  <div class="col-sm-4">
  
  ![A memorial to an Rock Island, Illinois, soldier who died in World War I after saving other soldiers in a mustard gas attack, quoting John 15:13. From the *Rock Island Argus* (Rock Island, Ill.), [14 Nov. 1919](http://chroniclingamerica.loc.gov/lccn/sn92053934/1919-11-14/ed-1/seq-7/#words=greater+love+hath+man+man+lay+down+life+friends).](img/greater-love.png)
  
  </div>
</div>

### How to use this site

There are three things that you can do on this site. First, you can use the [exploratory graphic](exploration.html) to investigate the trends in how different verses were quoted, and to find links back to *Chronicling America* for each quotations. Second, the [topics and verses](topics.html) page offers several mini-essays on specific topics or verses. And finally the [sources and methods](methods.html) section explains how I found the quotations. There you can also find links to the code I wrote and a [download of all the quotations that I found](methods.html#code-and-data).

### Suggested citation

If you use this project for academic work, here is a suggested citation.

> Lincoln Mullen, *America's Public Bible: Biblical Quotations in U.S. Newspapers*, website, code, and datasets (2016): &lt;<http://americaspublicbible.org>&gt;.


### Project news and roadmap

June 15, 2016

:   Initial release of the website and data

These are the goals for the project's future.

- Extend the model to other English language versions of the Bible (American Standard Version, Douay-Rheims version, Apocrypha)
- Extent the model to non-English language versions of the Bible (German Luther Bible)
- Extend the model to the Book of Mormon
- Improve model accuracy with additional training data from users of this website
- Time series analysis on the trends for verses, including seasonality and clustering
