---
title: "Biblical Passages Frequently Quoted Together"
---

Often more than one verse appears on a single newspaper page. A sermon, for instance, would quote multiple texts, or a political statement might marshall several passages in evidence.^[It is likely that some of these co-occurences also happen because the model for finding quotations has difficulty distinguishing between verses that are nearly identical, though I have removed many duplicates. See the [methods section](methods.html) for a fuller explanation.] These multiple quotations can be thought of as a network, where the passages that are quoted together on the same page are connected to one another. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      comment = NA, eval = FALSE)

library(dplyr)
library(stringr)
library(igraph)
library(purrr)
library(readr)
library(ggraph)
library(RColorBrewer)
```

```{r prepare-network-graph, include=FALSE, eval=FALSE}
quotations <- readRDS("_data/quotations-clean.rds") %>%
  mutate(reference = str_replace(reference, " \\(KJV\\)", ""))

books_of_bible <- read_csv("_data/books-of-bible.csv")

chapters <- quotations %>%
  mutate(reference = str_replace_all(reference, ":.+", ""))

chapter_counts <- chapters %>%
  count(reference)

edges_bi <- chapters %>%
  select(page, reference)

include_min <- 350

bi <- graph_from_data_frame(edges_bi, directed = FALSE)
V(bi)$type <- bipartite_mapping(bi)$type
all_verses_g <- bi %>% bipartite.projection(which = "true")
all_verses_g <- delete_edges(all_verses_g,
                             which(E(all_verses_g)$weight <= include_min))
separate_l <- decompose.graph(all_verses_g, min.vertices = 5)
big_g <- separate_l[[1]]
edges <- as_data_frame(big_g, "edges") %>% tbl_df()
nodes <- as_data_frame(big_g, "vertices") %>% tbl_df()

nodes <- nodes %>%
  mutate(book = str_replace(name, "\\s\\d+$", "")) %>%
  left_join(books_of_bible, by = "book") %>%
  left_join(chapter_counts, by = c("name" = "reference"))

g <- graph_from_data_frame(edges, directed = FALSE, nodes)
```

The visualization below shows the relationship between r V(g) %>% length() chapters in the Bible. For a chapter to be included, it had to have been quoted on the same page as another chapter at least r include_min times in the *Chronicling America* newspapers.

```{r chapters-network-graph, fig.cap="Chapters in the Bible most frequently quoted together", fig.width=12, fig.height=8, eval=FALSE}
set.seed(278)
ggraph(g, "igraph", algorithm = "nicely") +
  geom_edge_fan(aes(edge_width = weight,
                    alpha = weight)) +
  geom_node_point(aes(colour = type, size = n)) +
  scale_edge_width("Mutual appearances", range = c(1, 5), guide = "none") +
  scale_edge_alpha(range = c(0.2, 0.9), guide = "none") +
  # scale_color_brewer(palette = "Dark2", type = "qual") +
  scale_color_discrete("") +
  scale_size(range = c(2, 12), guide = "none") +
  ggforce::theme_no_axes() +
  geom_node_text(aes(label = name)) +
  labs(title = "Biblical passages frequently quoted together") +
  theme(legend.position = "bottom",
        panel.border = element_blank())
```
